Data Management and Domain Adjusting Strategies for Implementing Parallel SPH Method in JPUE Simulation

### 2016-01-04 Paper for submission.tex
add weak scalability

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE). Background grid is used to reduce neighbors searching cost and decompose (redecompose) the domain. Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid. As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.  
%to distinguish particles added at the same position but different simulation time step. 
Both particle and background grid are managed by hashtables which ensures quick and flexible accessing, adding and deleting of element. Additional link list attached to the hashtable is used to handle hash confliction. A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance. Furtherly, computational domain is adjusted during simulation to reduce computational cost. 
%Numerical tests show that our code has linear speed up.
These strategies can be further applied to many other implementations of meshfree methods, especially those implemetations that require flexible accessing, adding and deleting of particles and backgroud grids.
\end{abstract}

@@

Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing. (eg.\cite{ferrari2009new, crespo2015dualsphysics}). Less attention has been paid to developing of more flexible data management schemes for more complicated problems. 

@@

In both SPHysics\cite{dominguez2011optimization} and DualSPHysics\cite{crespo2015dualsphysics}, The physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells. This has two advantages: 1)access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known. But adding, deleting and especially accessing of particles are cumbersome. Angela Ferrari\cite{ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation. Storage problems caused by fix-size arrays are thereby also eliminated. We define C++ classes which contains all data of particle and bucket. As for the storage of data, we adopt an hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list. Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier. The SFC based numberring strategy is further extended to include time step information so that particles added at the same position but different time will have a different identifier. As for domain decomposion, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming. So we adopt an easy-programming scheme based on SFC \cite{patra1999efficient}.\\

@@

\caption{Basic Work flow for SPH}

@@

We will describe in this section these strategies which allows flexible deleting and adding.

@@

The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbor searching and domain decomposition. Both are defined as classes in C++. Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(process that the particle belongs to), primitive variables (variables show up as unknows in governing equations, like density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg. pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is vector of particle keys in our application). Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (27 neighbor buckets including itself for three dimension and 9 for two dimension) and possessed particles. Objects defined based on these two classes are then accesse through hash tables.

@@

The hash table has O(1) data accessing, adding and deleting properties when there is no confliction. How many conflictions will happen depends on both distribution of keys in the key space and size of the hash table. As the distribution of keys is determined by particles' initial location (or centroids of buckets), the hash table size is under our control. We can use very large hash table to minimize hash confliction on the expense of sparse data distribution which will lead to high cache missing and low memory efficiency. Or oppositely, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflictions. Abani\cite{patra2003data} did numerical experiments to make trade-off to determine hash table size.\\

@@

\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, these new particles will be stored in external link list}

@@

One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as higher digit and $t$ as lower digit of the large number. However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area. Only ghost particles for eruption boundary condition will be successsively added at the same place: bottom of the vent. That is to say, particles are distributed ununiformly in the $[0,1]^n \times \textbf{T}$ space as shown in figure \ref{fig:Particle_adding_with_link}. To avoid ununiform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}). 
\subsection{Load balance strategy}

@@

Particles used in the test problem can be categorized into four types based the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle. Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}). As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data. Instead of simply using number of contained particles as work load for bucket, work load of each bucket is determined by summing up work load weight of all particles within the bucket. The SFC sequence passing through centroids of all buckets now becomes a weighted sequence. Domain decomposition will conducted based on the weighted SFC of buckets.

@@

  \caption{Computational Cost for Different Steps Per Particle}

@@

Domain decomposition will be conducted based the weighted SFC of buckets. Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets. The particles are automatically split into several groups along with buckets that contain them. As SFC of buckets is a curve in the three dimensional space, partition of this curve will automatically lead to 3D domain decomposition. A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Dinesh\cite{kumar2013parallel}. A scalability comparision of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger. 

@@

Movement of particles, adding of new particles, adjusting of domain will lead to important load imbalance between processes. To handle this, computational load is monitored at given interval (The interval is optimized based on numerical experiments). Repartitioning is carried out if load imbalance is larger than a given tolenrance. \\
%Figure \ref{fig:t0_domain_decomposition} and \ref{fig:t10_domain_decomposition} show the initial decomposition and domain decomposion at 10 seconds.\\

%We adopt the same SFC based partition strategy as described by Patra \ref{patra1999efficient}.

%\begin{figure}[h]
%\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: initial domain decomposition}
%\label{fig:t0_domain_decomposition}
%\includegraphics[width=0.5\textwidth]{t0_domain_decomposition}
%\begin{figure}[h]
%\caption{An example of sub-dividing domain into 4 partitions using SFC Partitioning: decomposition at t=10s}
%\label{fig:t10_domain_decomposition}
%\includegraphics[width=0.5\textwidth]{t10_domain_decomposition}

@@

As a Lagrangian method, SPH is able to automatically adjust computational domain as the postition of the dicretization points are updated at every time step. However, for JPUE simulation, where some fluid ejects into stationary fluid and get mixed due to turbulence, the domain-adjusting feature of SPH will gone. Because the whole domain, which occupied by stationary fluid before eject fluids reaching there, has to be discretized at the very beginning of simulation. A lot of CPU time will be spent on computing of "stationary" particles. It is pure wasting of computational resources. If simulating of stationary particls can be avoided, the computational cost will be reduced greatly. To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. We propose a simple strategy to add such feature in our code with low computational cost. We add a scan function to monitor the most outside layer of the domain. When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition. The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such feature (see figure \ref{fig:Work_flow_adjust}).\\
\begin{figure}[h]

@@

\includegraphics[width=0.5\textwidth]{Work_flow_adjust}

@@

The start point of adding adjusting feature is adding a flag, which we refer to as involved-flag, into particle class. Particles are categorized into three groups based on the value of the involved-flag: involved (involved-flag=2), potential involved (involved-flag=1), and not involved(involved-flag=0). The involved particles are particles that have already been involved in eject mixing. The potential involved particles are particles that have not been involved in eject mixing but adjecent to involved particles. So they will be involved in the mixing process in the near future. These not involved particles are particles that far away from the eject source and are still at the initial state. For these not involved particles, it is meaningless to update its physical properties. As a consequence, searching for neighbors and data communitation for these particles are also unnecessary. That is to say, only potential involved and involved particles need to be simulated.
As simulation goes, the ejected fluid will reach larger area and more and more particles will be influenced. When originally stationary air is influenced by erupted material, the mass fraction of the erupted material of mixture will increase from zero to a positive value. So we can determine whether a particle is involved or not is based on whether the mass fraction of that particle is larger than a given threshold (10e-5 in our simulation). Other physical properties, such velocity, can also serve as alternative "swich criteria".\\

@@

A has-involved-flag is added to bucket class, too. Buckets are also categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0). Bucket that has any involved particle will be set to be has involved (has-involved-flag=3). And its neighbor buckets that do not contain any involved particle will be set to be has potential involved. 

@@

This domain adjusting process is shown in figure \ref{fig:Involved_initial} and \ref{fig:Involved_domain_adjusting}.

@@

Based on Navier-Stokes equations and several simplification assumption, the governing equations in Eulerian form are:

@@

Velocity is zero for non-slip wall boundary. We assume the boundary to be adiabatic and the heat flux is zero on the bounday. The pressure of the surrounding atmosphere is specified on pressure boundaries. Except for the pressure, boundary condtions for density, velocity, and energy will depend on the solution. We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Involved_initial} and figure \ref{fig:Involved_domain_adjusting}). 

@@

Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0, 4000]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2. The computational work load of test case 2 is 8 times of that of the test case 1. The computation was run for 20s physical time. The running time and speed up are shows in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}. The results shows linear speed up when number of processes is smaller than 16. The overhead of strong scalability can be increased by increasing total amount of work load which is implied by the fact that test case2 has better speed up than test case1 when number of processes is larger than 16. We also compared the performance of 2D domain decomposition and 3D domain decomposition. As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better strong scalability than 3D domain decomposition. When number processes increase, 3D domain decomposition shows a better speed up. The total number of footprint buckets in x-y plane is 64 for the 2D and 3D domain decomposition test case. When we request 128 processes, 64 processes will be totally idle at the beginning because there is only 64 buckets initially on x-y plane. As simulation goes, the domain will expand and more and more idle processes will start getting tasks. That's why we still get some speed up when number of process is 128.

@@

We developed data management strategies for parallel implementation of SPH method using MPI standard to simulate complicated problems, such as JUPE, which requires flexible and fast data retrievalling, adding and deleting. Neighbors searching and domain decomposition is based on background grid which overlaps the domain and keep stationary during simulation. SFC based index scheme, which provides a global numberring methodology that is purely coordinates dependent, is adopted to give each bucket an unique identifier. A time dependent key which is also based on SFC is used as identifier for particles. 
Hashtables with external link list are adopted for accessing  particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval. The code was further improved to 4 times faster by adjusting computational domain according to progress of simulation. 
Performance tests on our code shew a linear speed up and an overhead cast by total among of work load.

@@

The authors would like to thank...

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
This paper presents a parallel implementation of smoothed particle hydrodynamics (SPH) method using the message passing interface (MPI) standard to simulate a turbulent jet or plume which is ejected from a nozzle into a uniform environment(JPUE). Background grid is used to reduce neighbors searching cost and decompose (redecompose) the domain. Space Filing Curve(SFC) based index is adopted to assign an unique identifier to each background grid. As simulation of JUPE requires adding of new particles during simulation, time-dependent SFC based indexes are assigned to particles to guarantee uniqueness of identifiers.  
%to distinguish particles added at the same position but different simulation time step. 
Both particle and background grid are managed by hashtables which ensure quick and flexible accessing, adding and deleting of element. Additional link list attached to the hashtable is used to handle hash confliction. A SFC based three dimensional domain decomposition and a dynamic load balancing strategy are implemented to ensure good load balance. In addition, computational domain is adjusted during simulation to reduce computational cost. 
Numerical tests show that our code has acceptable strong scalability and good weak scalability.
These strategies can be further applied to many other implementations of meshfree methods, especially those implemetations that require flexible accessing, adding and deleting of particles and backgroud grids.
\end{abstract}

@@

Among existing CPU parallel SPH schemes, most of them focus on neighbors searching algorithm and dynamic load balancing. (eg.\cite{ferrari2009new, crespo2015dualsphysics}). Less attention has been paid to developing of more flexible data management schemes for more complicated problems. 

@@

In both SPHysics\cite{dominguez2011optimization} and DualSPHysics\cite{crespo2015dualsphysics}, The physical quantities of each particle (position, velocity, density…) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells. This has two advantages: 1)access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known. But adding, deleting and especially accessing of particles are cumbersome. Angela Ferrari\cite{ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation. Storage problems caused by fix-size arrays are thereby also eliminated. We define C++ classes which contains all data of particle and bucket. As for the management of data, we adopt an hash table to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding element, but also quicker access compared with linked list. Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier. The SFC based numberring strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers. As for domain decomposion, even though more complicated graph-based partition tools\cite{biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming. So we adopt an easy-programming scheme based on SFC \cite{patra1999efficient}.\\

@@

\caption{Basic work flow for SPH}

@@

We will describe in this section these strategies which satisfies these demand.

@@

The most basic data structure of SPH are particle, for problem discription, and bucket, for neighbors searching and domain decomposition. Both are defined as classes in C++. Infomation that contained in particle class can be categorized into six categorise: ID(the key), affiliate(rank of the process that the particle belongs to), primitive variables (variables show up as unknows in governing equations, eg. density, velocity, energy), secondary variables (properties that can be computed from primitive variables, they are stored to avoid repeatedly computing, eg. pressure, temperature ect.), flags (indicators, such as indicator for ghost particle and real particles, indicator for particles of different phases ect.) and neighbor infomation (it is a vector of particle keys in our application). Similarly,  Infomation that contained in bucket class can also be categorized into different categorise: ID(the key), affiliate(rank of the process that the bucket belong to), domain information (maximum and mimnum coordinates, boundary infomation), flags (indicators, such as indicator for guest and non-guest, indicator for active and inactive), neighbor infomation (keys of 27 neighbor buckets for three dimension and keys of 9 neighbor buckets for two dimension including its own key) and possessed particles. Objects defined based on these two classes are then accesse through hash tables.

@@

The hash table has O(1) data accessing, adding and deleting properties when there is no confliction. How many conflictions will happen depends on both distribution of keys in the key space and size of the hash table. As the distribution of keys is determined by particles' initial locations (or centroids of buckets), the hash table size is under our control. We can use very large hash table to minimize hash confliction on the expense of sparse data distribution which will lead to high cache missing and low memory efficiency. Or oppositely, we can use smaller hash table size to obtain high memory efficiency on the expense of having more hash conflictions. Abani\cite{patra2003data} did numerical experiments to examine
the effect of the table size on the dierent data management operations.\\

@@

\caption{Ununiform distribution of particles in the $[0,1]^n \times \textbf{T}$ space due to adding of new particles at a small portition of the domain, pointer to these new particles will be stored in external link list.}

@@

One natural way to hash time-dependent SFC based key $[k,t]$ is to convert the two elements in the key into one number taking $k$ as the higher digit and $t$ as the lower digit of the large number. However, for JPUE simulation, even though ghost particles for wall boundary condition and pressure boundary condition also need to be added during simulation, places for adding of these two types of ghost particles are previsouly empty area. Only ghost particles for eruption boundary condition will be successsively added at the same place: bottom of the vent. That is to say, particles are distributed ununiformly in the $[0,1]^n \times \textbf{T}$ space as shown in figure \ref{fig:Particle_adding_with_link}. To avoid ununiform, very sparse hash table and conserve locality of SFC, we only plug the first number, $k$, of the key, $[k,t]$, into the hashing function, equation (\ref{eq:hash_function}). 
\subsection{Load balancing strategy}

@@

Particles used in the test problem can be categorized into four types based the particle-type-flag: real particle, wall ghost particle, pressure ghost particle and eruption ghost particle. Ghost particles are for imposing of corresponding boundary conditions (see figure \ref{fig:Involved_initial}). As different types of particles involve different amount of computational work, shown by table \ref{tab:Computational_cost_particles} and table \ref{tab:Computational_cost_steps}, we assign different work load weight for different types of particles based on profilling data. Instead of simply using number of contained particles as work load for bucket, work load of each bucket is determined by summing up work load weight of all particles within the bucket. The SFC sequence passing through centroids of all buckets now becomes a weighted sequence. Domain decomposition will conducted based on the weighted SFC of buckets.\\

@@

  \caption{Computational Cost Per Particle for Different Steps}

@@

Domain decomposition will be conducted based the weighted SFC of buckets. Figure \ref{fig:SFC_domain_decomposition} shows how domain is decomposed based on partition of SFC of buckets. The particles are automatically split into several groups along with buckets that contain them. As SFC of buckets is a curve in the three dimensional space, partition of this curve will automatically lead to 3D domain decomposition. A 2D domain decomposition based on SFC of footprint buckets projected by three dimensional buckets was adopted by Dinesh\cite{kumar2013parallel}. A comparision of the scalability of these two schemes (see figure \ref{fig:2D_vs_3D_efficiency}) confirms that 3D domain decomposition is a better choice when processes number is larger. 

@@

Movement of particles, adding of new particles, adjusting of domain will lead to important load imbalance between processes. To handle this, computational load is monitored at a given interval (The interval is optimized based on numerical experiments). And repartitioning is carried out when load imbalance is larger than a given tolerance.

@@

As a Lagrangian method, SPH is able to automatically adjust computational domain as the postition of the dicretization points are updated at every time step. However, for JPUE simulation, where some fluid ejects into stationary fluid and get mixed due to turbulence, the domain-adjusting feature of SPH will gone. Because the whole domain, which occupied by stationary fluid before ejected fluid reaching there, has to be discretized at the very beginning of simulation. A lot of CPU time will be spent on computing of "stationary" particles. It is pure wasting of computational resources. If simulating of stationary particls can be avoided, the computational cost will be reduced greatly. To the best of the author's knowledge, no implementation of SPH has the feature of adjusting computational domain based on simulation. We propose a simple strategy to add such feature in our code with low computational cost. We add a scan function to monitor the most outside layer of the domain. When the ejected fluid reaches the boundary of the current domain, ghost particles (for pressure boundary condition) will be turn to real particles and then add new ghost particles for pressure boundary condition. The original work flow (see figure \ref{fig:Work_flow}) is modified to enable such feature (see figure \ref{fig:Work_flow_adjust}).\\
\includegraphics[scale=0.33]{Work_flow_adjust}

@@

@@

The start point of adding adjusting feature is adding a flag, which we refer to as involved-flag, into particle class. Particles are categorized into three groups based on the value of the involved-flag: involved (involved-flag=2), potential involved (involved-flag=1), and not involved(involved-flag=0). The involved particles are particles that have already been involved in eject mixing. The potential involved particles are particles that have not been involved in eject mixing but adjecent to involved particles. So they will be involved in the near future. These not involved particles are particles that far away from the eject source and are still at the initial state. For these not involved particles, it is meaningless to update its physical properties. As a consequence, searching for neighbors and data communitation for these particles are also unnecessary. That is to say, only potential involved and involved particles need to be simulated.
As simulation goes, the ejected fluid will reach larger area and more and more particles will be influenced. When originally stationary air is influenced by erupted material, the mass fraction of the erupted material will increase from zero to a positive value. So we can determine whether a particle is involved or not based on whether the mass fraction of that particle is larger than a given threshold (10e-5 in our simulation). Other physical properties, such velocity, can also serve as alternative "swich criteria".\\

@@

A has-involved-flag is added to bucket class, too. Buckets are then categorized into 3 types based on particles they contain: has involved (has-involved-flag=3), has potential involved(has-involved-flag=1) and has no involved (has-involved-flag=0). Bucket that has any involved particle will be set to be has involved (has-involved-flag=3). And its neighbor buckets that do not contain any involved particle will be set to be has potential involved. 

@@

This domain adjusting process is shown in figure \ref{fig:Domain_adjusting}.

@@

Based on Navier-Stokes equations and several simplifications, the governing equations in Eulerian form are:

@@

Velocity is zero for non-slip wall boundary. We assume the boundary to be adiabatic and the heat flux is zero on the bounday. The pressure of the surrounding atmosphere is specified on pressure boundaries. Except for the pressure, density, velocity, and energy will depend on the solution. We adpot ghost particles to impose these different type of boundary conditions (see figure \ref{fig:Domain_adjusting}). 

@@

\includegraphics[scale=0.33]{weak_loglog}
\caption{Weak scalability test results}
\label{fig:weak_loglog}
%Experiments have been carried out on the computational cluster of Center for Computational Research (CCR) at Buffalo. 
The initial domain is $[-4800m,4800m]\times [-4800m,4800m] \times [0m, 6000m]$, with smoothing length (we set initial intervals between particles equal to smoothing length) equals to 200m and 100m respectively for test case1 and test case2. The computational work load of test case 2 is 8 times of that of the test case 1. The simulations run for 20s physical time. The consumed time and speed up are shown in figure \ref{fig:2cases_time} and \ref{fig:2cases_efficiency}. Linear speed up is observed when number of processes is smaller than 16. Test case2 shows better speed up than test case1 which implies that the overhead of strong scalability can be increased by increasing total amount of work load. We also compared the performance of 2D domain decomposition and 3D domain decomposition. As shown by figure \ref{fig:2D_vs_3D_efficiency}, 2D domain decomposition shows a little bit better speed up than 3D domain decomposition for 32 processese. This can be explained by the fact that 2D domian decomposition can get more regular subdomain and as a results, need less communication. When number of processes increase, it becomes harder for 2D decomposition to balance workload, that is why 3D domain decomposition has better speed up for processes number equals to 64 and 128. The weak scalability test is conducted with the same initial domain and various smoothing length. Each simulation runs for 400 loops. The average number of real particles of each process keep constant. As shown in figure \ref{fig:weak_loglog}, simulation time increases pretty slowly when number of processes is larger than 132. the relative sharp increase in simulation time from 32 processes to 132 processes might cause by a increase of number of neighbor processes. 
%That can be explained by the fact that the total number of footprint buckets in x-y plane is 64 in the test case.
%When we request 128 processes, 64 processes will be totally idle at the beginning. As simulation goes, the domain will expand and more and more idle processes will start getting tasks. That's why we still get some speed up when number of process is 128. 3D domain decomposition is able to divide the work load in a more flexible way and achieve better strong scalability.

@@

We developed data management strategies for parallel implementation of SPH method using MPI standard to simulate complicated problems, such as JUPE, which requires flexible and fast data retrievalling, adding and deleting. Neighbors searching and domain decomposition is based on background grid which overlaps the domain and keep stationary during simulation. SFC based index scheme, which provides a global numberring methodology which is purely coordinates dependent, is adopted to give each bucket an unique identifier. A time dependent key which is also based on SFC is used as identifier for particle. 
Hashtables with external link list are adopted for accessing particles and buckets data. Based on weighted particle work load, a dynamic load balance strategy is developed by checking load balance and redecomposing the domain at an optimize interval. The performance of the code was further improved to several times faster by adjusting computational domain according to progress of simulation. 
Scalability tests on our code shew acceptable strong scalability and good weak scalability. 3D domain decomposition provided better strong scalability than 2D domain decomposition when number of processes is larger.

@@

Computational results reported here were performed at
the Center for Computational Research at the University at
Buffalo. This project is supported by Grants No. NSF
1131074 from the National Science
Foundation.
